Checkerboards And Missing Species Combinations: Are Ecological Communities

### 2016-01-27 ManuscriptMarkDown.Rmd
Corrected heat maps and table

---
title: 'Checkerboards And Missing Species Combinations: Are Ecological Communities
  Assembled By Chance?'
author: "Nicholas J. Gotelli"
date: "24 January 2016"
output:
  pdf_document: default
  html_document:
    highlight: haddock
    keep_md: yes
    theme: cerulean
  word_document: default
---

@@

@@

In one of his assembly rules, Diamond proposed that some pairs of species were such strong competitors that they never occured together on the same island, forming a "checkerboard distribution". For example, from the matrix above, we notice that the species in rows 3 and 6 (x and y) form a checkerboard pair and never co-occur:

>>>>

---
title: 'Checkerboards And Missing Species Combinations: Are Ecological Communities
  Assembled By Chance?'
author: "Nicholas J. Gotelli"
date: "24 January 2016"
output:
  html_document:
    highlight: tango
    theme: cerulean
  pdf_document: default
  word_document: default
---

@@

```{r echo=FALSE}
library(EcoSimR)
library(ggplot2)
library(reshape2)
library(knitr)
set.seed(2016)
m <- melt(dataWiFinches)
m <- as.data.frame(cbind(rep(1:17,19),rep(1:19,each=17),m$value))

p <- ggplot( m, aes(x=V2, y=V1, fill=V3))
p + geom_tile() + theme(legend.position="none") + xlab("Islands") + ylab("Species")
```

@@

In one of his assembly rules, Diamond proposed that some pairs of species were such strong competitors that they never occured together on the same island, forming a "checkerboard distribution". For example, from the matrix above, we notice that the species in rows 8 and 10 (*Loxigilla noxis* and *Tiara olivacea*) form a checkerboard pair and never co-occur:

```{r echo=FALSE}
m <- melt(dataWiFinches[c(8,10),])
m <- as.data.frame(cbind(rep(1:2,19),rep(1:19,each=2),m$value))

p <- ggplot( m, aes(x=V2, y=V1, fill=V3))
p + geom_tile() + theme(legend.position="none") + xlab("Islands") + ylab("Species")

```

The data can be organized into a two-way contingency table:

```{r echo=FALSE}
Mychecker <- dataWiFinches[c(8,10),-1]
Mychecker <- as.matrix(Mychecker)
MyCon <- matrix(0,nrow=2,ncol=2, dimnames = list(c("Species A Present","Species A Absent"),c("Species B Present","Species B Absent")))
MyCon[1,1] <- sum(Mychecker[1,]==1 & Mychecker[2,]==1)
MyCon[1,2] <- sum(Mychecker[1,]==1 & Mychecker[2,]==0)
MyCon[2,1] <- sum(Mychecker[1,]==0 & Mychecker[2,]==1)
MyCon[2,2] <- sum(Mychecker[1,]==0 & Mychecker[2,]==0)
kable(MyCon, padding=0)
```

and a Fisher's exact test yields (
